---
title: "Adaptation Impact"
author: "Megan Patricia Rogers"
date: "6/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description
This Markdown file documents the construction of data that characterizes adaptation
potential of cultivar choice and soil organic matter management. It should be noted
that due to a lack of solid data, evaluating the significance of these findings is not possible at this point. This part of the paper's data analysis is more explicitly speculative and was included to draw attention to the need to further explore and predict the impacts of adaptive agricultural practices such as better soil management and bio-engineering. 

#### Package loading:
``` {r, warning=FALSE, message=FALSE}
library(tidyverse) # suite of packages for data manipulation/visualization
library(Hmisc) # more data analysis
library(dplyr)  # data fame manipulation
library(plotrix)  # makes calculating standard error easier
library(skimr) # for creating summaries
library(magrittr) # enable piping
```

#### Load data:
```{r}
processed_adaptation_potential <- read.csv("processed_adaptation_potential.csv", 
                                           header = TRUE, sep = ",", dec = ".")
```

#### Creating New Dataframe 
##### Calculating USA Statistics
```{r}

# Create new data frame with averages that you can compare

# USA
usa_adaptation_potential <- processed_adaptation_potential %>%
  filter(country == "United States") # use USA data frame to calculate mean adaptation potential 

# Tillage
usa_tillage <- usa_adaptation_potential %>% 
  filter(adaptation_type == "Tillage")

usa_till_potential_percentage =
  mean(usa_tillage$adaptation_potential_percentage)

usa_till_potential_se =
  std.error(usa_tillage$adaptation_potential_percentage)

# Cultivar
usa_cultivar <- usa_adaptation_potential %>% 
  filter(adaptation_type == "Cultivar")

usa_cultivar_potential_percentage =
  mean(usa_cultivar$adaptation_potential_percentage)

usa_cultivar_percentage_se =
  std.error(usa_cultivar$adaptation_potential_percentage)
```


##### Calculating Canada Statistics
``` {r}
# Canada
canada_adaptation_potential <- processed_adaptation_potential %>%
  filter(country == "Canada")

# Tillage
canada_tillage <- canada_adaptation_potential %>% 
  filter(adaptation_type == "Tillage")

canada_till_potential_percentage =
  mean(canada_tillage$adaptation_potential_percentage)

canada_till_potential_se =
  std.error(canada_tillage$adaptation_potential_percentage)

# Cultivar
canada_cultivar <- canada_adaptation_potential %>% 
  filter(adaptation_type == "Cultivar")

canada_cultivar_potential_percentage =
  mean(canada_cultivar$adaptation_potential_percentage)

canada_cultivar_potential_se =
  std.error(canada_cultivar$adaptation_potential_percentage)

```

#### Combining USA and Canada Averages 
``` {r}
adaptation_potential_by_country = 
  data.frame(country = c("United States", "Canada"),
             tillage_percent_potential = c(usa_till_potential_percentage, canada_till_potential_percentage),
             tillage_percent_se = c(usa_till_potential_se, canada_till_potential_se),
             cultivar_percent_potential = c(usa_cultivar_potential_percentage, canada_cultivar_potential_percentage),
             cultivar_percent_se = c(usa_cultivar_percentage_se, canada_cultivar_potential_se))
             

```

#### Bar Plot of Tillage Potential
```{r, warning=FALSE}

ggplot(adaptation_potential_by_country, aes(x=country, y=tillage_percent_potential, 
                          color=country)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = tillage_percent_potential - tillage_percent_se,
                    ymax = tillage_percent_potential + tillage_percent_se), width = .6) +
  labs(title="Projected Yield Increase due to Tillage Practices Per Country", 
       x="Country", 
       y="Yield Increase (%)")

```
#### Bar Plot of Cultivar Potential
```{r}
ggplot(adaptation_potential_by_country, aes(x=country, y=cultivar_percent_potential, 
                          color=country)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = cultivar_percent_potential - cultivar_percent_se,
                    ymax = cultivar_percent_potential + cultivar_percent_se), width = .6) +
  labs(title="Projected Yield Increase due to Cultivar Improvements Per Country", 
       x="Country", 
       y="Yield Increase (%)")
```

#### Exploring Statistical Significance
``` {r}

# tillage:
hist(canada_cultivar$adaptation_potential_percentage,
     main = "Histogram of estimated cultivar improvements Canada",
     xlab = "Estimated percent yield change")

hist(usa_cultivar$adaptation_potential_percentage,
     main = "Histogram of estimated cultivar improvements USA",
     xlab = "Estimated percent yield change")

# cultivar:
hist(usa_tillage$adaptation_potential_percentage,
     main = "Histogram of estimated tillage improvements USA",
     xlab = "Estimated percent yield change")


```
