---
title: "Data Processing and Wrangling"
author: "Megan Patricia Rogers"
date: "6/15/2022"
output: html_document
---

### Description
This Markdown file documents the processing of two datasheets to filter for and 
rename relevant variables to construct more managable dataframes for subsequent analysis. This paper uses data, provided by Hasegawa et al., that is split between two 
data sets as described below:

##### 1. Projected Impacts
includes data on climate and yield projections for multiple climate change 
scenarios on 6 continents

##### 2. Adaptation Potential
calculated differences in yield projections between adapation/no adaptation 
scenarios between scenarios in the same study. directly used data from
Projected Impacts


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Package loading:
``` {r, warning=FALSE, message=FALSE}
library(tidyverse)  # suite of packages for data manipulation/visualization
library(Hmisc)  # more data analysis
library(dplyr)  # data frame manipulation
library(skimr)  # for creating summaries
library(magrittr)  # enable piping
```

#### Load data from set working directory:
documentation source: 
[Link] (https://stat.ethz.ch/R-manual/R-devel/library/utils/html/read.table.html)
```{r}

# read CSV files into data frames:

projected_impacts <- read.csv("2022_06_Projected_Impacts_Big_Data.csv", 
                                 header = TRUE, sep = ",", dec = ".")

adaptation_potential <- read.csv("2022_06_Adaptation_Potential_Big_Data.csv", 
                                 header = TRUE, sep = ",", dec = ".")
```

#### Data processing for projected_impacts -> processed_projected_impacts dataframe
The table below summarizes the variables included in the processed data set. The included script details the process of selecting and renaming variables for ease of analysis. 

```{r, include=TRUE}

# read projected_impacts metadata file into data frame
projected_impacts_variable_metadata <- 
  read.csv("2022_06_projected_impacts_variable_metadata.csv", 
                                 header = TRUE, sep = ",", dec = ".")
# create table from data frame
dplyr::glimpse(projected_impacts_variable_metadata) %>%
  knitr::kable()  # just to show the variables included.

# how do you just include the nice table and not the janky one? 

```


```{r, include=TRUE}
processed_projected_impacts <- projected_impacts %>%
  dplyr::select("ID", 
                crop = "Crop", 
                country = "Country", 
                sublocation = "Site.location.", 
                continent = "Region", 
                current_avg_temp = 
                  "Current.Average.Temperature_point_coordinate..dC.",
                current_annual_precipitation = 
                  "Current.Annual.Precipitation...mm.._point_coordinate",
                mid_point_year = "Future_Mid.point", 
                time_slice = "Time.slice", 
                climate_scenario = "Climate.scenario",
                local_temp_change_since_2005 = "Local.delta.T.from.2005", 
                local_precipitation_change_since_2005 = 
                  "Annual.Precipitation.change..from.2005..mm.",
                global_temp_chance_since_2005 = "Global.delta.T.from.2005", 
                projected_yield = "Projected.yield..t.ha.",
                climate_yield_change_from_2005 = "Climate.impacts.relative.to.2005",
                yield_change_per_C = "Climate.impacts.per.dC....",
                yield_change_per_decade = "Climate.impacts.per.decade....",
                ) %>%
  dplyr::filter(crop == "Maize") %>%  # interested in corn
  dplyr::filter(climate_scenario == "RCP4.5") %>%  # filter for RCP4.5-based predictions
  dplyr::filter(continent == "North America")  # focus on North America
```


#### Data processing for adaptation_potential -> processed_adaptation_potential dataframe
The table below summarizes the variables included in the processed data set. The included script details the process of selecting and renaming variables for ease of analysis. 

```{r, include=TRUE}
# read projected_impacts metadata file into data frame
adaptation_potential_variable_metadata <- 
  read.csv("2022_06_adaptation_potential_variable_metadata.csv", 
                                 header = TRUE, sep = ",", dec = ".")
 # create table from data frame
dplyr::glimpse(adaptation_potential_variable_metadata) %>% 
  knitr::kable()  # just to show the variables included.

```


```{r, include=TRUE}
processed_adaptation_potential <- adaptation_potential %>%
  dplyr::select("ID", 
                crop = "Crop", 
                country = "Country", 
                sublocation = "Site.location.", 
                continent = "Region", 
                mid_point_year = "Future_Mid.point", 
                time_slice = "Time.slice", 
                climate_scenario = "Climate.scenario",
                scale = "Scale",
                crop = "Crop",
                sublocation = "Site.location.",
                continent = "Region",
                adaptation_type = "Adaptation.type",
                soil_organic_matter = "Soil.organic.matter.management",
                tillage = "Tillage",
                cultivar = "Cultivar",
                adaptation_potential_actual = "Adaptation.potential_actual.yield",
                adaptation_potential_percentage = "Adaptation.potential_..."
                ) %>%
  dplyr::filter(crop == "Maize") %>%  # interested in corn
  dplyr::filter(climate_scenario == "RCP4.5") %>%  # filter for RCP4.5-based predictions
  dplyr::filter(continent == "North America")  # focus on North America
```

#### Export the two dataframes to working directory folder
``` {r}
write.csv(processed_projected_impacts,
          "/Users/meganrogers/Desktop/Plutorial/Final_Project_Directory/processed_projected_impacts.csv", 
          row.names = FALSE)

write.csv(processed_adaptation_potential,
          "/Users/meganrogers/Desktop/Plutorial/Final_Project_Directory/processed_adaptation_potential.csv", 
          row.names = FALSE)
```








